Using Bayesian Regression to Quantify Sexual Dimorphism in Dinosaurs

### 2024-02-13 paperAndPoster/paper.tex
feat: updated graphs with default save options

\title{Using Bayesian Regression to Quantify Sexual Dimorphism in Extinct Taxa}
\author{Eric Campbell}
\begin{document}
\maketitle

@@

\section{Methods}
The modification to the method in \cite{saittaEffectSizeStatistical2020} is presented below.

@@

Despite these advantages, there remain many area in which this method could be improved. The most critical, as noted in \cite{saittaEffectSizeStatistical2020}, is sex determination. As shown by Figure \ref{fig:alligatorSexPredictionError} sex determination accuracy was a strong predictor of the error in the level of dimorphism. There are two complementary paths which should be taken towards alleviating this issue. The first is the incorporation of the uncertainty in sex determination into the posterior curves generated. At the moment we behave as though the sex that we predict is correct, with no uncertainty, when the reality is that this is a significant source of error. This would result in widening the posterior distributions for the levels of dimorphism, but would not address the underlying issue. For that, we should look at alternate means of predicting sex, such as cluster analysis involving traits other than the one that we are looking at.

@@

For all data and analyses used here, see \url{https://github.com/EricRobertCampbell/aps-2024-poster-sexual-dimorphism}.

\printbibliography[title=Bibliography]

>>>>

\title{Using Bayesian Regression to Quantify Sexual Dimorphism in Dinosaurs}
\author{Eric Campbell}
\begin{document}
\maketitle

@@

\section{Methods}
Bayesian regression is quite different from classical regression. To estimate a parameter, there needs to be a \emph{prior} representing the information that you have before seeing the data. You can then calculate the \emph{likelihood} of the data that you observed for the different parameter values. The optimal combination of the prior and the likelihood is given by Bayes' Theorem and is termed the \emph{posterior}. In general, the result of a Bayesian analysis will be a probability distribution, not a single number \parencite{mcelreathStatisticalRethinkingBayesian2020}. An illustration of this method is shown in Figure \ref{fig:bayesianExplanation}.

\begin{figure}[H]
	\centering
	\includegraphics[width = \textwidth]{images/bayesianExample.png}
	\caption{To estimate human height, we start with what we know before seeing any data --- the \emph{prior}. Then a sample of heights is collected and the \emph{likelihood} of that sample for each of the parameter values is calculated. The combination of those, our final belief about the value of the parameter, is the \emph{posterior} distribution.}
	\label{fig:bayesianExplanation}
\end{figure}

The modification to the method in \cite{saittaEffectSizeStatistical2020} is presented below.

@@

Despite these advantages, there remain many areas in which this method could be improved. The most critical, as noted in \cite{saittaEffectSizeStatistical2020}, is sex determination. As shown by Figure \ref{fig:alligatorSexPredictionError} sex determination accuracy was a strong predictor of the error in the level of dimorphism. There are two complementary paths which should be taken towards alleviating this issue. The first is the incorporation of the uncertainty in sex determination into the posterior curves generated. At the moment we behave as though the sex that we predict is correct, with no uncertainty, when the reality is that this is a significant source of error. This would result in widening the posterior distributions for the levels of dimorphism, but would not address the underlying issue. For that, we should look at alternate means of predicting sex, such as cluster analysis involving traits other than the one that we are looking at.

@@

\printbibliography[title=Bibliography]
For all data and code used here, along with additional analyses, see \url{https://github.com/EricRobertCampbell/aps-2024-poster-sexual-dimorphism}.
